\subsection*{Abbreviations}
SCN: suprachiasmatic nucleus; $\textrm{MSF}_\textrm{sc}$: mid sleep time on free days after correcting for sleep debt during week days; FDR: false discovery rate; MAE: median absolute error; GEO: Gene Expression Omnibus; SPC: sparse principal component.

\subsection*{Acknowledgments}
AG Herzel, AG Kramer \todo{Anyone else to be acknowledged?}

\subsection*{Conflict of interest}
\todo{Please fill in which authors are working for Beiersdorf and if there are any conflicts of interest} 
%Is any author working for Beiersdorf?, \textcolor{red}{which markets skin care products}. The remaining authors declare that they have no competing interests.\todo{Please fill in which authors are working for Beiersdorf and if there are any conflicts of interest} 


\subsection*{Data accessibility}
The gene expression microarray data of dermis and epidermis generated in this study are available at the Gene Expression Omnibus (GEO) database and can be accessed under the accession number \textcolor{red}{GSEblabla}. The source code is available through GitHub: \textcolor{red}{link github}. 

\subsection*{Funding}
\todo{Please add/fill in funding agencies}
%\textcolor{red}{Beiersdorf paid for 100\% of the costs of clinical work reported in this paper. This work is also supported by } 

\subsection*{Author contributions}
Study design and conceptualization: BA, MdO, AK, DK and HH. Experimental Methodology: SK, AG, JdZ and \textcolor{blue}{CN}. Bioinformatic methodology: MdO, BA and KJ. Investigation: BA, MdO, KJ, AK and HH. Writing (original draft and editing): MdO and BA. Writing (review): MdO, BA, AK, HH and JdZ. Funding acquisition: BA, AK and HH. \todo{Please double check and change author contributions}

\subsection*{Author affiliations}
\textsuperscript{1}Institute for Theoretical Biology -- Laboratory of Theoretical Chronobiology, Humboldt Universit\"at zu Berlin and Charit\'e Universit\"atsmedizin Berlin, Philippstra\ss e 13, House 4, 10115 Berlin, Germany. \textsuperscript{2}Institute for Medical Immunology -- Laboratory of Chronobiology, Charité Universitätsmedizin Berlin, Charitéplatz 1, 10117 Berlin, Germany. \textsuperscript{3}Berlin Institute of Health -- Center for Regenerative Therapies (BCRT), Charit\'e Universit\"atsmedizin Berlin, Augustenburger Platz 1,
13353 Berlin, Germany. \textsuperscript{4}Institute of Physiology -- Sleep Research \& Clinical Chronobiology, Charit\'e Universit\"atsmedizin Berlin, Charit\'eplatz 1, 10117 Berlin, Germany.\todo{please add/check affiliations}